Manuscript_Github_Pchabaudi_pir

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Reading in the TPM to the figure file
PCA just RMT

```{r times_stages_labels}
#Use the rMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

@@

```{r pca_calculation, results = TRUE}
twentyfour_tpm <- twentyfour_ct_comb
rownames_geneid <- twentyfour_tpm$Geneid
twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
rownames(twentyfour_tpm_forpca) <- rownames_geneid
twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
#Need to look into what exactly scale and center are doing
```

@@

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

@@

(
  plot <- ggplot(twentyfour_tpm_pca_coords,
                 aes(
                   x = PC1,
                   y = PC2,
                   col = time,
                   shape = transmission
                 )) +
    geom_point(size = 3,
               alpha = 0.8) +
    scale_color_manual(values = colornames) +
    scale_shape_manual(
      values = setNames(
        c(17,19),
        unique(twentyfour_tpm_pca_coords$transmission)
      )
    ) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(4, 'mm'),
          legend.text = element_text(size = 7))
)

@@

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

@@

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

@@

cir_plot_order <- cir_plot %>% 
  filter(time == '14h',
         transmission == 'rMT') %>% 
  arrange(desc(tpm))

@@

           y <- ggplot(
             data = cir_plot_filter,
             mapping = aes(
               x = Geneid,
               y = tpm
             )
           ) +
             stat_summary(fun.y = median, 
                          fun.ymin = median, 
                          fun.ymax = median,
                          geom = "bar",
                          fill = 'lightblue',
                          width = 0.75, 
                          size = 0.25, 
                          position = 'dodge') +
             geom_point(stat = 'identity', 
                        position = position_dodge(width = 0.75),
                        size = point_size) +
             scale_y_continuous(breaks = breaks,
                                limits = limits) +
             facet_grid(.~subfam, 
                        scales = "free_x",
                        space = "free_x",
                        labeller = labeller(subfam = function(string) {
                          str_remove_all(string, 
                                         pattern = '(rMT|SBP)_')
                        })) +
             theme_classic() +
             ylab('log(TPM + 1)')+
             ggtitle(paste(unique(cir_plot_filter$time)))

@@

           strip_t <- which(grepl('strip-t', g$layout$name))
           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
           fills <- case_when(str_detect(subfam_vec, 
                                         pattern = 'a') ~ clade_col['ancestral'],
                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                              str_detect(subfam_vec, 
                                         pattern = 'ref') ~ 'white')
           k <- 1
           for (i in strip_t) {
             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
             k <- k+1
           }

@@

```{r SorL_barchart}
tpm_data_SorL <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm))

@@

```{r clades_prop_barchart}
tpm_data_clade <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  # mutate(stage =  factor(stage, 
  #                        levels = ordering(stage, 
  #                                          order_vector))) %>% 
  group_by(transmission,
           stage, 
           clade) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  arrange(stage)

@@

```{r SorL_barchart_transmission-facet}
tpm_data_SorL <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  mutate(SorL = factor(SorL, 
                       levels = sort(unique(SorL))))

@@

tpm_data_twentyfour <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)],
         chrom = str_extract(Geneid, 
                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
         ) %>% 
  mutate(
    SorL = str_extract(clade, 
                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
    # stage =  factor(stage, 
    #                 levels = ordering(stage, order_vector))
  ) %>% 
  group_by(transmission,
           stage, 
           Geneid,
           clade,
           chrom,
           SorL) %>% 
  dplyr::summarise(tpm = mean(tpm))

@@

gene_expression_df <- twentyfour_ct_comb%>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(pir_gene = ifelse(Geneid %in% cir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, 
           pir_gene,
           transmission) %>% 
  slice_sample(n = length(cir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))

>>>>

```{r times_stages_labels}
#Use the RMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r read_TPM}
#Need to use the calculateTPM script first.

twentyfour_rMT <- read_csv('Pchabaudi_TPM_allBioReps.csv')

twentyfour_avgtpm <- list(
  'RMT' = list(
    tpm = read_csv('Pchabaudi_TPM_average.csv'),
    experiment = 'RMT'
  )
)

@@

```{r pca_calculation, results = TRUE}
twentyfour_tpm <- twentyfour_rMT
rownames_geneid <- twentyfour_tpm$Geneid
twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
rownames(twentyfour_tpm_forpca) <- rownames_geneid
twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
#Need to look into what exactly scale and center are doing
```

@@

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,1])

@@

(
  plot <- ggplot(twentyfour_tpm_pca_coords,
                 aes(
                   x = PC1,
                   y = PC2,
                   col = time
                 )) +
    geom_point(size = 3,
               alpha = 0.8) +
    scale_color_manual(values = colornames) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(4, 'mm'),
          legend.text = element_text(size = 7))
)

@@

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

@@

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

@@

cir_plot_order <- cir_plot %>% 
  filter(time == '14h',
         transmission == 'RMT') %>% 
  arrange(desc(tpm))

@@

           y <- ggplot(
             data = cir_plot_filter,
             mapping = aes(
               x = Geneid,
               y = tpm
             )
           ) +
             stat_summary(fun.y = median, 
                          fun.ymin = median, 
                          fun.ymax = median,
                          geom = "bar",
                          fill = 'lightblue',
                          width = 0.75, 
                          size = 0.25, 
                          position = 'dodge') +
             geom_point(stat = 'identity', 
                        position = position_dodge(width = 0.75),
                        size = point_size) +
             scale_y_continuous(breaks = breaks,
                                limits = limits) +
             facet_grid(.~subfam, 
                        scales = "free_x",
                        space = "free_x",
                        labeller = labeller(subfam = function(string) {
                          str_remove_all(string, 
                                         pattern = '(RMT|SBP)_')
                        })) +
             theme_classic() +
             ylab('log(TPM + 1)')+
             ggtitle(paste(unique(cir_plot_filter$time)))

@@

           strip_t <- which(grepl('strip-t', g$layout$name))
           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(RMT|SBP)_')
           fills <- case_when(str_detect(subfam_vec, 
                                         pattern = 'a') ~ clade_col['ancestral'],
                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                              str_detect(subfam_vec, 
                                         pattern = 'ref') ~ 'white')
           k <- 1
           for (i in strip_t) {
             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
             k <- k+1
           }

@@

```{r SorL_barchart}
tpm_data_SorL <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm))

@@

```{r clades_prop_barchart}
tpm_data_clade <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  # mutate(stage =  factor(stage, 
  #                        levels = ordering(stage, 
  #                                          order_vector))) %>% 
  group_by(transmission,
           stage, 
           clade) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  arrange(stage)

@@

```{r SorL_barchart_transmission-facet}
tpm_data_SorL <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  mutate(SorL = factor(SorL, 
                       levels = sort(unique(SorL))))

@@

tpm_data_twentyfour <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)],
         chrom = str_extract(Geneid, 
                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
         ) %>% 
  mutate(
    SorL = str_extract(clade, 
                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
    # stage =  factor(stage, 
    #                 levels = ordering(stage, order_vector))
  ) %>% 
  group_by(transmission,
           stage, 
           Geneid,
           clade,
           chrom,
           SorL) %>% 
  dplyr::summarise(tpm = mean(tpm))

@@

gene_expression_df <- twentyfour_ct_comb%>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(pir_gene = ifelse(Geneid %in% cir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, 
           pir_gene,
           transmission) %>% 
  slice_sample(n = length(cir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))
